# Methodology {#methods}

This chapter introduces the V-Dem dataset and discuss the methodological background of `tectr`'s construction.

## V-Dem

### Introduction to the database

The Varieties of Democracy Institute is concerned with measuring different aspects of democracy. It distinguishes between seven high-level principles: electoral, liberal, participatory, deliberative, egalitarian, majoritarian and consensual. These are measured by a variable in the interval $[0,1]$ and each consists of several mid- and low-level indices. The low-level indices are coded with the help of several country experts who answer a detailed questionnaire. Most questions can be answered by an ordinal scale of five alternatives. Consider, as an example, the variable "Disclosure of campaign donations":

> **Are there disclosure requirements for donations to national election campaigns?**
>
> 0: No. There are no disclosure requirements.
>
> 1: Not really. There are some, possibly partial, disclosure requirements in place but they are not observed or enforced most of the time.
>
> 2: Ambiguous. There are disclosure requirements in place, but it is unclear to what extent they are observed or enforced.
>
> 3: Mostly. The disclosure requirements may not be fully comprehensive (some donations not covered), but most existing arrangements are observed and enforced.
>
> 4: Yes. There are comprehensive requirements and they are observed and enforced almost all the time.

The answers are then analyzed for inter-coder reliability and a standardized average of the responses together with a confidence interval which contains 68 % of the probability mass is created. Lower-level indices are created from these answers which are summarized in mid-level and then high-level indices. An overview over the structure can be found in appendix D of the codebook [@vdem-codebook2018]. The database contains data on 201 countries between 1789 and 2017. [@vdem2018, @Pemstein2018]

### `vdem.tectr`

I have created the package `vdem.tectr` which contains the country-year dataset version 8. It can be downloaded via [github](github.com/sflippl/vdem.tectr):

```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("sflippl/vdem.tectr")
```

The package contains three datasets:

* `df_vdem`: This dataset contains all variables from the varieties of democracy dataset where interval variables are numeric and categorical variables are saved as factors or ordered factors where appropriate.
* `vdem_spatial`: This *simple features* object [@sf] contains the polygon shapes of the different countries for every year between 1945 and 2017. The borders in 2017 can be seen in \@ref(fig:vdem-spatial). I have used the CShapes dataset [@Weidmann2010, @Weidmann2010a], sovereignty- and state-level maps data from [Natural Earth](www.naturalearthdata.com) and the details from the document on country coding units from V-Dem [@Coppedge2018]. Note that the coded country borders by V-Dem do not constitute an endorsement of controversial entities such as Zanzibar. The function `vdem_geocode` makes it possible to join the shapes polygons to a data frame as long as the columns `country_name` and `year` exist.

```{r vdem-spatial, fig.cap="Country borders in 2017 in the V-Dem database", echo = FALSE}
vdem_spatial %>% 
  filter(end_year == 2017) %>% 
  ggplot() + 
  geom_sf() + 
  theme_map()
```

* `vdem` which contains the variables from `df_vdem`, the country shapes from `vdem_spatial` and further metainformation (see [below](#application))

Details on these datasets and the reproducible code can be found in the folder "data-raw" in the package.

## Package construction

The package has been constructed with the packages `devtools` [@devtools], `roxygen2` [@roxygen2] and `testthat` [@testthat]. The Bachelor's Thesis has been written with `bookdown` [@bookdown-1, @bookdown-2].