# Application {#application}

This chapter is concerned with the application of `tectr` to the V-Dem database. At first, it will present the metaframe and discuss the `fx_write` function. The second section will present `fx_info` and `fx_output` whereas the third section will consider `fx_ggplot`.

## The metaframe

The folder "data-raw" contains the creation of the metaframe in detail. I have imported the codebook [@vdem-codebook2018] with the help of the `pdftools` package [@pdftools]. Subsequently, I have read out different categories of information about the variables. The resulting data frame has been exported by `fx_write_json` which yielded the data frame `mf_revisable`:

```{r}
mf_revisable
```

In particular, the different information categories have been collapsed into a filepath called "main".

The files themselves have the following structure:

```
[
  {
    "fxInfo_name": "Electoral democracy index",
    "fxInfo_notes": "",
    "fxInfo_ordering": "",
    "fxInfo_project_manager": ["Jan Teorell"],
    "fxInfo_question": "To what extent is the ideal of electoral democracy in its fullest sense achieved?",
    "fxInfo_responses": [],
    "fxInfo_scale": "Interval, from low to high (0-1)",
    "fxInfo_sources": "v2x_freexp_altinf v2x_frassoc_thick v2x_suffr v2xel_frefair v2x_elecoff",
    "name": "v2x_polyarchy"
  }
]
```

I have only included a few exemplary fields in order to demonstrate that such a file is well-readable by humans.

At this stage, these variables may be modified by hand in order to add additional fields or correct typos.

Afterwards, the metaframe is completed by adding the parameters of `fxGeom`. This results in the `vdem`-object:

```{r}
metaframe(vdem)
```

## `fx_info` 

We can use `vdem` to produce uncomplicated and well-structured descriptive tables, for instance table \@ref(tab:desc).

```{r desc}
vdem %>% 
  select(v2x_polyarchy, v2x_libdem, v2x_partipdem, v2x_delibdem, v2x_egaldem) %>% 
  fx_info(c("name", "question", "stats"), 
          statistics = list(
            mean = function(x, ...) mean(x, na.rm = TRUE),
            quantile = function(x, ...) 
              quantile(x, probs = seq(0.25, 0.75, 0.25), 
                       na.rm = TRUE)
            )) %>% 
  select(-name) %>% 
  fx_output("table", out_format = "markdown")
```

A more vertical structure is given below by the "collapse"-output-form:

```{r results = 'asis'}
vdem %>% 
  select(v2x_polyarchy) %>% 
  fx_info(c("name", "question", "clarification", "stats"), 
          statistics = list(
            mean = function(x, ...) mean(x, na.rm = TRUE),
            quantile = function(x, ...) quantile(x, na.rm = TRUE)
            )
          ) %>% 
  fx_output("collapse", 
            cell_scheme = "*{name}*\n{value}", 
            cell_sep = "\n\n",
            out_format = "markdown")
```

## `fx_ggplot`

As the general features of `fx_ggplot` have already been presented in the previous chapter, I will focus on a few changes due to the adaptations of the metaframe. Firstly, I have provided a consistent set of palettes, as can be seen in figure \@ref(fig:consistent).

```{r consistent, fig.cap = "The ordinal scale provides a discretized version of the original scale.", out.width = "60%", fig.align="center"}
p1 <- vdem %>% 
  filter(year == 2015) %>% 
  select(v2clrelig_osp, country_name, year) %>% 
  vdem_geocode() %>% 
  fx_ggplot(aes(geom = geometry, fill = v2clrelig_osp)) + 
  my_theme

p2 <- vdem %>% 
  filter(year == 2015) %>% 
  select(v2clrelig_ord, country_name, year) %>% 
  vdem_geocode() %>% 
  fx_ggplot(aes(geom = geometry, fill = v2clrelig_ord)) + 
  my_theme

p1/p2
```

Furthermore, we plot a time series with confidence intervals for different countries in figure \@ref(fig:time).

```{r time, fig.cap = "A time series of the electoral democracy index for Germany, France and Belgium", out.width = "50%", fig.align="center"}
vdem %>% 
  filter(country_name %in% c("Germany", "France", "Belgium")) %>% 
  select(country_name, year, contains("v2x_polyarchy")) %>% 
  fx_ggplot(aes(x = year, y = v2x_polyarchy, colour = country_name)) +
  my_theme
```

If we include all countries, `fx_ggplot` prevents overplotting, as in figure \@ref(fig:time-all).

```{r time-all, fig.cap = "A time series of the electoral democracy index for all countries", out.width = "50%", fig.align="center"}
vdem %>% 
  select(country_name, year, contains("v2x_polyarchy")) %>% 
  fx_ggplot(aes(x = year, y = v2x_polyarchy)) +
  my_theme
```

Of course, we can also plot two democracy indices against each other, see figures \@ref(fig:scatter) and \@ref(fig:scatter-facet).

```{r scatter, fig.cap = "The electoral democracy index and the liberal democracy index.", out.width = "50%", fig.align="center"}
vdem %>% 
  select(v2x_polyarchy, v2x_libdem) %>% 
  fx_ggplot(aes(x = v2x_polyarchy, y = v2x_libdem)) +
  my_theme
```

```{r scatter-facet, fig.cap = "Electoral and liberal democracy index in Germany, France and Belgium over time", out.width = "50%", fig.align="center"}
vdem %>% 
  filter(country_name %in% c("Germany", "France", "Belgium")) %>% 
  select(v2x_polyarchy, v2x_libdem, country_name, year) %>% 
  fx_ggplot(aes(x = v2x_polyarchy, y = v2x_libdem,
                colour = year), 
            facet_vars = vars(country_name)) +
  my_theme
```

The time series work for the low-level indices, as well, as can be seen in figure \@ref(fig:low).

```{r low, fig.cap = "The electoral democracy index and the liberal democracy index.", out.width = "50%", fig.align="center"}
vdem %>% 
  filter(country_name %in% c("Germany", "France", "Belgium")) %>% 
  select(contains("v2meharjrn"), country_name, year) %>% 
  fx_ggplot(aes(x = year, y = v2meharjrn_osp, colour = country_name)) + 
  my_theme
```

Figure \@ref(fig:low-disc) shows the discretized version.

```{r low-disc, fig.cap = "Harassment of journalists in Germany, France and Belgium over time (discretized)", out.width = "50%", fig.align="center"}
vdem %>% 
  filter(country_name %in% c("Germany", "France", "Belgium")) %>% 
  select(contains("v2meharjrn"), country_name, year) %>% 
  fx_ggplot(aes(x = year, y = v2meharjrn_ord, colour = country_name)) +
  my_theme
```

Finally, we plot a discretized variable against a high-level index in figure \@ref(fig:disc-high).

```{r disc-high, fig.cap = "The electoral democracy index and the liberal democracy index.", out.width = "50%", fig.align="center"}
vdem %>% 
  select(contains("v2meharjrn"), v2x_polyarchy, country_name, year) %>% 
  fx_ggplot(aes(x = v2meharjrn_ord, y = v2x_polyarchy)) +
  my_theme
```

As a final proof of concept, [appendix B](#vardescs) contains an automated report of the high-level indices.